### 2012-07-13 doc/latex/hhmm/hhmm.max-min-table.tex
Work on journal article

\begin{tabular} {| c || r | r || r | r || r | r || r | r |}
\hline
%\multirow{2}{*}{Split} 
Split Size & \multicolumn{2}{|c||}{\fontspec{TeX Gyre Pagella}\textsc{Naive-Split}} & \multicolumn{2}{|c||}{\fontspec{TeX Gyre Pagella}\textsc{Histogram-Split}} & \multicolumn{2}{|c||}{\fontspec{TeX Gyre Pagella}\textsc{Words-Split}} & \multicolumn{2}{|c|}{\fontspec{TeX Gyre Pagella}\textsc{Times-Split}} \\ \cline{2-9}
($n$) & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c||}{Max} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c||}{Max} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c||}{Max} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Max} \\ \hline
2 & {365.7} & {408.0 } & {376.3} & {397.5 } & {386.1} & {387.6 } & \textit{386.9} & \fontspec{TeX Gyre Pagella}\textbf{\textit{386.9 }} \\ \hline
4 & {176.1} & {214.4 } & {186.8} & {205.0 } & {184.6} & {200.9 } & \textit{193.4} & \fontspec{TeX Gyre Pagella}\textbf{\textit{193.4 }} \\ \hline
8 & {84.6} & {122.8 } & {84.8} & {107.6 } & {88.4} & {101.9 } & \textit{96.7} & \fontspec{TeX Gyre Pagella}\textbf{\textit{96.7 }} \\ \hline
16 & {40.8} & {70.8 } & {40.5} & {62.0 } & {45.3} & {55.0 } & \textit{48.4} & \fontspec{TeX Gyre Pagella}\textbf{\textit{48.4 }} \\ \hline
32 & {19.2} & {38.3 } & {18.8} & {37.8 } & {20.5} & {29.2 } & \textit{24.2} & \fontspec{TeX Gyre Pagella}\textbf{\textit{24.2 }} \\ \hline
64 & {9.2} & {26.1 } & {9.2} & {25.4 } & {9.4} & {16.7 } & \textit{12.1} & \fontspec{TeX Gyre Pagella}\textbf{\textit{12.1 }} \\ \hline
128 & {2.7} & {16.6 } & {4.1} & {13.0 } & {3.7} & {9.3 } & {5.9} & \fontspec{TeX Gyre Pagella}\textbf{6.2 } \\ \hline
256 & {0.7} & {10.2 } & {1.3} & {8.0 } & {1.4} & {6.6 } & {2.9} & \fontspec{TeX Gyre Pagella}\textbf{4.6 } \\ \hline
512 & {0.0} & {6.9 } & {0.3} & {6.3 } & {0.6} & \fontspec{TeX Gyre Pagella}\textbf{4.6 } & {1.1} & \fontspec{TeX Gyre Pagella}\textbf{4.6 } \\ \hline
\end{tabular}